---
title: "chapter5"
author: Yingjia
date: "2023-12-04"
---



# chapter 5: Dimensionality reduction techniques


```{r}
date()
```

## load data

```{r}
library(readr)
human <- read_csv("https://raw.githubusercontent.com/KimmoVehkalahti/Helsinki-Open-Data-Science/master/datasets/human2.csv")
dim(human)
```


## move the country names to rownames

```{r}

library(tibble)
human_ <- column_to_rownames(human, "Country")
summary(human_)
```

```{r}
#graphical overview of the data 
library(ggplot2)
library(GGally)
ggpairs(human_, progress = FALSE)
```
There is higher positive correlations between Ado.Birth and Mat.Mor, Edu.Exp and Edu2.FM. 

There is higher negative correlations between Life.Exp and Mat.Mor, Edu.Exp and Mat.Mor, Edu2.FM and Mat.Mor, Ado.Birth and Life.Exp, Ado.Birth and Edu.Exp.


## PCA

```{r}
pca_human <- prcomp(human_)
biplot(pca_human, choices = 1:2, xlabs=rep(".", nrow(human_)))
```

## Standardize the variables and PCA

```{r}
human_scale <- scale(human_)
pca_human_scale <- prcomp(human_scale)
biplot(pca_human_scale, choices = 1:2, xlabs=rep(".", nrow(human_scale)))
```
```{r}
s <- summary(pca_human)
pca_pr <- round(1*s$importance[2, ], digits = 5)
pca_pr

```


```{r}
s_scale <- summary(pca_human_scale)
pca_pr_scale <- round(1*s_scale$importance[2, ], digits = 5)
pca_pr_scale
```

## tea data
```{r}
tea <- read.csv("https://raw.githubusercontent.com/KimmoVehkalahti/Helsinki-Open-Data-Science/master/datasets/tea.csv", stringsAsFactors = TRUE)
#View(tea)
head(tea)
#dim(tea)
```

```{r}
library(tidyr)
library(dplyr)
keep_columns <- c("Tea", "How", "how", "sugar", "where", "lunch")
tea_time <- select(tea, keep_columns)
pivot_longer(tea_time, cols = everything()) %>% 
  ggplot(aes(value)) + facet_wrap("name", scales = "free")
```
```{r}
#MCA
#install.packages('FactoMineR')
library(FactoMineR)
mca <- MCA(tea_time, graph = FALSE)
plot(mca, invisible=c("ind"), graph.type = "classic")

```

