# chapter 2: Regression and model validation


*Describe the work you have done this week and summarize your learning.*

- Describe your work and results clearly. 
- Assume the reader has an introductory course level understanding of writing and reading R code as well as statistical methods.
- Assume the reader has no previous knowledge of your data or the more advanced methods you are using.

```{r}
date()
```

Here we go again...

## Load the students2014 data

```{r}
df = read.csv('./data/learning2014.txt', sep = ',', stringsAsFactors = F, header = T)
dim(df)
```
```{r}
head(df)
```

```{r}
str(df)
```

The students2014 data contains 166 observations and 7 variables (gender, age, attitude, deep, stra, surf, points).

## A graphical overview of the data

```{r}
par(mfrow = c(2,4))
barplot(table(df$gender), main = 'gender')
hist(df$age)
hist(df$attitude)
hist(df$deep)
hist(df$stra)
hist(df$surf)
hist(df$points)
```
```{r}
pairs(df[, -1])
```


##  Fit a regression model 
```{r}
lm.1 <- lm(points~attitude+deep+stra, data = df)
summary(lm.1)
```

remove the two variables (surf and stra) which did not have a statistically significant relationship with the target variable (exam points) and fit the regression model with the remaining variables

```{r}
lm.2 <- lm(points~attitude, data = df)
summary(lm.2)
```

## Explanations on the fitted model
R-squareds tell how large proportion of the variance in Y the line explains compared to the total variance in Y .
```{r}
summary(lm.2)$r.squared
```
In this case, the linear model on attitude explains around 20% of the variation in exam points and therefore leaves about 80% of variation in exam points as something that cannot be explained by (a linear effect of) attitude only. 


## Produce the following diagnostic plots

### Residuals vs Fitted values, Normal QQ-plot and Residuals vs Leverage
```{r}
par(mfrow = c(2,2)) 
plot(lm.2)
```

1. Residuals vs. fitted should not show a pattern where the distribution of residuals varies depending on the fitted values. Thus, is is good if the red line showing the mean of data points is close to a horizontal line.

2. QQ-plot should ideally be on the diagonal line, in which case the residuals are Normally distributed and SEs and P-values of the model coefficients are reliable. In this case, the QQ-plot deviated from the diagonal line.

3. Residuals vs. leverage should not show points that reach outside the curved areas of Cook’s distance 1. Otherwise, such points would have a high influence on the slope of the linear model, and hence the lm() estimates from such data would not be reliable.

